### 2022-01-07 code/R/Tables.Rmd
Added R script for tables

---
title: "Tables"
author: "Hjalte Mann"
date: "7/1/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("gt")
library(gt)
library(tidyverse)
library(glue)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:
```{r}

# Define the start and end dates for the data range
start_date <- "2010-06-07"
end_date <- "2010-06-14"

sp500

# Create a gt table based on preprocessed
# `sp500` table data
sp500 %>%
  filter(date >= start_date & date <= end_date) %>%
  select(-adj_close) %>%
  gt() %>%
  tab_header(
    title = "S&P 500",
    subtitle = glue("{start_date} to {end_date}")
  ) %>%
  fmt_date(
    columns = date,
    date_style = 3
  ) %>%
  fmt_currency(
    columns = c(open, high, low, close),
    currency = "USD"
  ) %>%
  fmt_number(
    columns = volume,
    suffixing = TRUE
  )
```
```{r}
ts = read.csv2("U:/BITCue/Projekter/TrackingFlowers/testResults/TrackingScores.csv")
print(ts)

```