### 2022-02-02 text/SupportManuscriptDraft.Rmd
Manuscript update Rev 2

---
new_session: FALSE
title: |
  **Supporting Information: Towards individual-based pollination ecology: Automatic tracking of life histories of individual flowers**
author:
- Hjalte M. R. Mann
- Alexandros Iosifidis
- Toke T. Høye
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  word_document: default
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
  html_document:
    df_print: paged
fontfamily: mathpazo
fontsize: 12pt
geometry: margin = 1in
header-includes:
- \usepackage{setspace}\doublespacing
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{dcolumn}
- \usepackage{caption}
- \usepackage{float}
- \usepackage{afterpage}
- \usepackage{siunitx}
- \usepackage{amsmath}
keywords: Tracking, flowering phenology, Arctic
bibliography: ./library.bib
csl: ./journal-of-ecology.csl
---

```{r eval = FALSE, echo=FALSE}
library(knitr)
knit('ManuscriptDraft.Rmd')
rmarkdown::render("SupportManuscriptDraft.Rmd", output_file = "SupportManuscriptDraft.docx")
system2("open","SupportManuscriptDraft.docx")
```

```{r echo=FALSE}
# Function for creating word comments in Rmarkdown
# Use: This text contains a `r word_comment("This is the comment", "comment.")`.

word_comment <- function(comment, highlight = "") {
  if (isTRUE(knitr:::pandoc_to() == "docx")) {
    paste0('[', comment, ']{.comment-start id="0" author="Hjalte M."',
           'date="2020-01"}', highlight, '[]{.comment-end id="0"}')
  }
}
```

\newpage

**Table S1:** Results of the filtering algorithm for filtering distances 5-500 with a stepsize of 50 for all three series.

![](../figures/FullFilteringDistanceTest.png){ width=100% }