### 2022-05-25 code/R/VisualizeTracks.Rmd
Added script for visualizing tracks to exemplify errors.

---
title: "VisualizeTracks"
author: "Hjalte Mann"
date: '2022-05-12'
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(stringr)
library(GGally)
library(plotly)
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
meta_2018 = read.csv("../../data/annotations/DetectionPhenology_2018_Meta_Clean_Prepped.csv")

meta_2018 = meta_2018 %>% 
  select(FileName, DOY, DateTime)

colnames(meta_2018) = c("filename", "DOY", "time")

```

```{r}
THUL01Anno = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2020_05_17_NorwayAnnotations_NARS-04_IndividualAnnotations_FRCNN_Metrics.csv", header = T)

THUL01Tracked = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/_parameterTest_NARS-04_3/parameterTest_NARS-04_maxDisap_10_runMean_20_maxDist_500.csv")

m = merge(THUL01Anno, THUL01Tracked, by = c("filename","x_min", "y_min", "x_max", "y_max"))

m = merge(m, meta_2018, by = "filename")

m %>% 
  filter(id_gt == "Flower_15" | id_gt == "Flower_28"| id_gt == "Flower_25") %>% 
ggplot() + 
  geom_point(aes(x = frame, y = id_gt, colour = as.factor(objectID)), size = 4, shape = 15)

mp = m %>% 
  filter(id_gt == "Flower_15" | id_gt == "Flower_28"| id_gt == "Flower_25") %>% 
  ggplot() +
   geom_line(mapping = aes(DOY,id_gt, group=id_gt, colour = as.factor(objectID)), size = 4) +
  ylab("") +
  ggtitle("Max gap = 10 Running mean = 1 Max distance = 0") +
   theme_bw() +
   theme(legend.position = "none",
         panel.grid.minor.y = element_blank(),
         panel.grid.major.y = element_blank(),
         panel.grid.minor.x = element_blank(),
         panel.grid.major.x = element_blank()) #,
         #text = element_text(size=30))#,
         #axis.text.y=element_blank(),
         #axis.ticks.y=element_blank(),
         #axis.text=element_text(colour="black"),
         #panel.spacing.y = unit(0.7, "lines"))# +
  #theme(plot.margin = margin(0.5,1,0.5,0.5, "cm"))

```

```{r}
THUL01Anno2 = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2020_05_17_NorwayAnnotations_NARS-04_IndividualAnnotations_FRCNN_Metrics.csv", header = T)

THUL01Tracked2 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/_parameterTest_NARS-04_3/parameterTest_NARS-04_maxDisap_10_runMean_1_maxDist_0.csv")

m2 = merge(THUL01Anno2, THUL01Tracked2, by = c("filename","x_min", "y_min", "x_max", "y_max"))

m2 = merge(m2, meta_2018, by = "filename")

m2 %>% 
  filter(id_gt == "Flower_15" | id_gt == "Flower_28"| id_gt == "Flower_25") %>% 
ggplot() + 
  geom_point(aes(x = frame, y = id_gt, colour = as.factor(objectID)), size = 4, shape = 15)

m2p = m2 %>% 
  filter(id_gt == "Flower_15" | id_gt == "Flower_28"| id_gt == "Flower_25") %>% 
  ggplot() +
   geom_line(mapping = aes(DOY,id_gt, group=id_gt, colour = as.factor(objectID)), size = 4) +
  ylab("") +
  ggtitle("Max gap = 10 Running mean = 1 Max distance = 0") +
   theme_bw() +
   theme(legend.position = "none",
         panel.grid.minor.y = element_blank(),
         panel.grid.major.y = element_blank(),
         panel.grid.minor.x = element_blank(),
         panel.grid.major.x = element_blank())#,
         #text = element_text(size=30))#,
         #axis.text.y=element_blank(),
         #axis.ticks.y=element_blank(),
         #axis.text=element_text(colour="black"),
         #panel.spacing.y = unit(0.7, "lines"))# +
  #theme(plot.margin = margin(0.5,1,0.5,0.5, "cm"))

```

```{r}

m$type = "Max gap = 10, Running mean = 20, Max distance = 500"
m2$type = "Max gap = 10, Running mean = 1, Max distance = 0"

m3 = bind_rows(m, m2)

m$c = "NA"

m3$c[m3$objectID == "17"] = "0"
m3$c[m3$objectID == "15"] = "0"

m3$c[m3$objectID == "13"] = "1"
m3$c[m3$objectID == "10"] = "2"

m3$c[m3$objectID == "12"] = "1"
m3$c[m3$objectID == "9"] = "2"

m3$c[m3$objectID == "27"] = "3"
m3$c[m3$objectID == "15"] = "0"

errorPlot = m3 %>% 
  filter(id_gt == "Flower_15" | id_gt == "Flower_28"| id_gt == "Flower_25") %>% 
  ggplot() +
   geom_line(mapping = aes(DOY,id_gt, group=id_gt, colour = c), size = 4) +
  ylab("Ground truth flower ID") +
  facet_wrap(vars(type), ncol = 1) +
  #ggtitle("B") +
   theme_bw() +
   theme(legend.position = "none",
         panel.grid.minor.y = element_blank(),
         panel.grid.major.y = element_blank(),
         panel.grid.minor.x = element_blank(),
         panel.grid.major.x = element_blank(),
         text = element_text(size=18))

errorPlot

ggsave("aplot.png", errorPlot, height = 4, width = 10, dpi = 800)

```

```{r}
#unique(m$objectID[m$id_gt == "Flower_32"])
#unique(m$id_gt[m$objectID == "9"])

#m %>% 
#ggplot() + 
#  geom_point(aes(x = x_c, y = y_c, colour = as.factor(id_gt)), size = 1)

#m %>% 
#ggplot() + 
#  geom_point(aes(x = x_c, y = y_c, colour = as.factor(objectID)), size = 1)

#m %>% 
  #group_by(id_gt, objectID) %>% 
  #mutate(gr = n()) %>% 
#ggplot() + 
#  geom_point(aes(x = frame, y = id_gt, colour = as.factor(objectID)), size = 4, shape = 15)

#m %>% 
# ggplot() + 
 #  geom_point(aes(x = frame, y = as.factor(objectID), colour = id_gt), size = 4, shape = 15)

#m %>% 
#  filter(id_gt == "Flower_15" | id_gt == "Flower_28"| id_gt == "Flower_25") %>% 
#ggplot() + 
#  geom_point(aes(x = frame, y = as.factor(objectID), colour = id_gt), size = 4, shape = #15)

```

```{r}
NARS13Anno = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv", header = T)

nars13points = NARS13Anno %>% 
  mutate(x_c = (x_min + x_max)/2, y_c = (y_min + y_max)/2) %>% 
ggplot() + 
  geom_point(aes(x = x_c, y = y_c, colour = as.factor(id_gt)), size = 1, shape = 4, stroke = 1) +
  labs(x = NULL, y = NULL, title = NULL) +
  scale_x_continuous(limits = c(0,6080), expand = c(0, 0)) + 
  scale_y_reverse(limits = c(3420, 0), expand = c(0, 0)) +
   theme_bw() +
   theme(legend.position = "none",
    axis.ticks.length=unit(0, "pt"),
    #panel.background = element_rect(fill = "transparent", colour = NA),
    #plot.background = element_rect(fill = "transparent", colour = NA),
    panel.grid = element_blank())#,
    #panel.border = element_blank())

nars13points
ggsave("nars13points.png", nars13points, width = 10, height = 5.64, dpi = 800)

```