### 2023-01-13 src/R/MOTMetrics.Rmd
Resubmission -first commit

---
title: "MOTMetrics"
author: "Hjalte Mann"
date: '2022-05-02'
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(stringr)
library(GGally)
library(plotly)
knitr::opts_chunk$set(echo = TRUE)
```

####### MOTMETRICS MMFix PARAMETER TEST ########

# NYAA-04
```{r fig.height=4}
NYAA04d1 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/NYAA-04_parameterTest_Evaluation_20220513.csv")

NYAA04_p1 = NYAA04d1 %>% 
  ggplot() +
  geom_point(aes(x=runMean, y = maxDisap, size = num_switches, colour = num_switches)) + 
  ggtitle("NYAA-04 Parameter combination") +
  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")

NYAA04_p2 = NYAA04d1 %>% 
  arrange(num_switches, ascending = T) %>% 
  slice_head(n = 10) %>% 
  ggplot() +
  geom_point(aes(x=runMean, y = maxDisap, size = num_switches, colour = num_switches)) + 
  ggtitle("NYAA-04 Ten Lowest Mismatches") +
  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")

NYAA04_p1
NYAA04_p2

ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NYAA-04_AllCombinations.png', NYAA04_p1)
ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NYAA-04_TenLowest.png', NYAA04_p2)

rm(NYAA04_p1)
rm(NYAA04_p2)
```

# NARS-13

```{r fig.height=4}
NARS13d1 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/NARS-13_parameterTest_Evaluation_20220513.csv")

NARS13_p1 = NARS13d1 %>%
  ggplot() +
  geom_point(aes(x=runMean, y = maxDisap, size = num_switches, colour = num_switches)) +
  ggtitle("NYAA-04 Parameter combination") +
  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")

NARS13_p2 = NARS13d1 %>%
  arrange(num_switches, ascending = T) %>%
  slice_head(n = 10) %>%
  ggplot() +
  geom_point(aes(x=runMean, y = maxDisap, size = num_switches, colour = num_switches)) +
  ggtitle("NYAA-04 Ten Lowest Mismatches") +
  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")

NARS13_p1
NARS13_p2

ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NARS-13_AllCombinations.png', NARS13_p1)
ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NARS-13_TenLowest.png', NARS13_p2)

rm(NARS13_p1)
rm(NARS13_p2)
```

# NARS-04
```{r fig.height=4}
NARS04d1 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/NARS-04_parameterTest_Evaluation_20220513.csv")

NARS04_p1 = NARS04d1 %>% 
  ggplot() +
  geom_point(aes(x=runMean, y = maxDisap, size = num_switches, colour = num_switches)) + 
  ggtitle("NYAA-04 Parameter combination") +
  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")

NARS04_p2 = NARS04d1 %>% 
  arrange(num_switches, ascending = T) %>% 
  slice_head(n = 10) %>% 
  ggplot() +
  geom_point(aes(x=runMean, y = maxDisap, size = num_switches, colour = num_switches)) + 
  ggtitle("NYAA-04 Ten Lowest Mismatches") +
  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")

NARS04_p1
NARS04_p2

ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NARS-04_AllCombinations.png', NARS04_p1)
ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NARS-04_TenLowest.png', NARS04_p2)

rm(NARS04_p1)
rm(NARS04_p2)
```

# NARS-17
```{r fig.height=4}
NARS17d1 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/NARS-17_parameterTest_Evaluation_20220513.csv")

NARS17_p1 = NARS17d1 %>% 
  ggplot() +
  geom_point(aes(x=runMean, y = maxDisap, size = num_switches, colour = num_switches)) + 
  ggtitle("NYAA-04 Parameter combination") +
  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")

NARS17_p2 = NARS17d1 %>% 
  arrange(num_switches, ascending = T) %>% 
  slice_head(n = 10) %>% 
  ggplot() +
  geom_point(aes(x=runMean, y = maxDisap, size = num_switches, colour = num_switches)) + 
  ggtitle("NYAA-04 Ten Lowest Mismatches") +
  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")

NARS17_p1
NARS17_p2

ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NARS-17_AllCombinations.png', NARS17_p1)
ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/NARS-17_TenLowest.png', NARS17_p2)

rm(NARS17_p1)
rm(NARS17_p2)
```

# THUL-01
```{r fig.height=4}
THUL01d1 = read.csv("U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/THUL-01_parameterTest_Evaluation_20220513.csv")

THUL01_p1 = THUL01d1 %>% 
  ggplot() +
  geom_point(aes(x=runMean, y = maxDisap, size = num_switches, colour = num_switches)) + 
  ggtitle("NYAA-04 Parameter combination") +
  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")

THUL01_p2 = THUL01d1 %>% 
  arrange(num_switches, ascending = T) %>% 
  slice_head(n = 10) %>% 
  ggplot() +
  geom_point(aes(x=runMean, y = maxDisap, size = num_switches, colour = num_switches)) + 
  ggtitle("NYAA-04 Ten Lowest Mismatches") +
  facet_wrap(vars(maxDist))+scale_color_gradient(low="blue", high="red")

THUL01_p1
THUL01_p2

ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/THUL-01_AllCombinations.png', THUL01_p1)
ggsave('U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/figures/THUL-01_TenLowest.png', THUL01_p2)

rm(THUL01_p1)
rm(THUL01_p2)
```

### Explore combinations

# NYAA-04
```{r echo = FALSE}
print("NYAA-04")
nyaa04_gt = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2021_12_13_NorwayAnnotations_NYAA-04_IndividualAnnotations_FRCNN_Metrics.csv")

print('True number of flowers')
length(unique(nyaa04_gt$id_gt))

# Minimum num_switches
print("Lowest number of mismatches")
min(NYAA04d1$num_switches)
NYAA04Minnum_switches = NYAA04d1[NYAA04d1$num_switches == min(NYAA04d1$num_switches),]

# All zero
NYAA04AllZero = NYAA04d1 %>% 
  filter(maxDisap == 0 & maxDist == 0 & runMean == 1)

# Correct number of flowers and minimum num_switches
NYAA04CorrectflowerMinnum_switches = NYAA04d1 %>% 
  filter(num_tracks == length(unique(nyaa04_gt$id_gt)))

NYAA04CorrectflowerMinnum_switches = NYAA04CorrectflowerMinnum_switches[NYAA04CorrectflowerMinnum_switches$num_switches == min(NYAA04CorrectflowerMinnum_switches$num_switches),]

```

```{r echo = FALSE}
print("NARS-13")
nars13_gt = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2020_04_30_NorwayAnnotations_NARS-13_IndividualAnnotations_FRCNN_Metrics.csv")

print('True number of flowers')
length(unique(nars13_gt$id_gt))

# Minimum num_switches
print("Lowest number of mismatches")
min(NARS13d1$num_switches)
NARS13Minnum_switches = NARS13d1[NARS13d1$num_switches == min(NARS13d1$num_switches),]

# All zero
NARS13AllZero = NARS13d1 %>%
  filter(maxDisap == 0 & maxDist == 0 & runMean == 1)

# Correct number of flowers and minimum num_switches
NARS13CorrectflowerMinnum_switches = NARS13d1 %>%
  filter(num_tracks == length(unique(nars13_gt$id_gt)))
NARS13CorrectflowerMinnum_switches = NARS13CorrectflowerMinnum_switches[NARS13CorrectflowerMinnum_switches$num_switches == min(NARS13CorrectflowerMinnum_switches$num_switches),]
```

# NARS-04
```{r echo = FALSE}
print("NARS-04")
nars04_gt = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2020_05_17_NorwayAnnotations_NARS-04_IndividualAnnotations_FRCNN_Metrics.csv")

print('True number of flowers')
length(unique(nars04_gt$id_gt))

# Minimum num_switches
print("Lowest number of mismatches")
min(NARS04d1$num_switches)
NARS04Minnum_switches = NARS04d1[NARS04d1$num_switches == min(NARS04d1$num_switches),]

# All zero
NARS04AllZero = NARS04d1 %>% 
  filter(maxDisap == 0 & maxDist == 0 & runMean == 1)

# Correct number of flowers and minimum num_switches
NARS04CorrectflowerMinnum_switches = NARS04d1 %>% 
  filter(num_tracks == length(unique(nars04_gt$id_gt)))
NARS04CorrectflowerMinnum_switches = NARS04CorrectflowerMinnum_switches[NARS04CorrectflowerMinnum_switches$num_switches == min(NARS04CorrectflowerMinnum_switches$num_switches),]
```

# NARS-17
```{r echo = FALSE}
print("NARS-17")
nars17_gt = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2021_12_30_NorwayAnnotations_NARS-17_IndividualAnnotations_FRCNN_Metrics.csv")

print('True number of flowers')
length(unique(nars17_gt$id_gt))

# Minimum num_switches
print("Lowest number of mismatches")
min(NARS17d1$num_switches)
NARS17Minnum_switches = NARS17d1[NARS17d1$num_switches == min(NARS17d1$num_switches),]

# All zero
NARS17AllZero = NARS17d1 %>% 
  filter(maxDisap == 0 & maxDist == 0 & runMean == 1)

# Correct number of flowers and minimum num_switches
NARS17CorrectflowerMinnum_switches = NARS17d1 %>% 
  filter(num_tracks == length(unique(nars17_gt$id_gt)))
NARS17CorrectflowerMinnum_switches = NARS17CorrectflowerMinnum_switches[NARS17CorrectflowerMinnum_switches$num_switches == min(NARS17CorrectflowerMinnum_switches$num_switches),]
```

# THUL-01
```{r echo = FALSE}
print("THUL-01")
thul01_gt = read.csv("U:/BITCue/Projekter/TrackingFlowers/data/annotations/2020_05_15_NorwayAnnotations_THUL-01_IndividualAnnotations_FRCNN_Metrics.csv")

print('True number of flowers')
length(unique(thul01_gt$id_gt))

# Minimum num_switches
print("Lowest number of mismatches")
min(THUL01d1$num_switches)
THUL01Minnum_switches = THUL01d1[THUL01d1$num_switches == min(THUL01d1$num_switches),]

# All zero
THUL01AllZero = THUL01d1 %>% 
  filter(maxDisap == 0 & maxDist == 0 & runMean == 1)

# Correct number of flowers and minimum num_switches
THUL01CorrectflowerMinnum_switches = THUL01d1 %>% 
  filter(num_tracks == length(unique(thul01_gt$id_gt)))
THUL01CorrectflowerMinnum_switches = THUL01CorrectflowerMinnum_switches[THUL01CorrectflowerMinnum_switches$num_switches == min(THUL01CorrectflowerMinnum_switches$num_switches),]
```

# Table dataframe

```{r}
trackScores = setNames(data.frame(matrix(ncol = 11, nrow = 0)), c('Site', 'Type', 'Flowers','Points', 'Max gap', 'Max distance', 'Running mean', 'Mismatches', 'MOTA', '>80% Accuracy','Tracks'))

# Add THUL-01 to table
# All zero
THUL01AllZero$Site = "2018 THUL B"
THUL01AllZero$Type = "PZ"

THUL01AllZeroTable = THUL01AllZero %>% 
  select(Site,Type,num_unique_objects, num_detections, maxDisap, maxDist, runMean,num_switches, mota, mostly_tracked, num_tracks)

trackScores[nrow(trackScores) + 1,] = c(THUL01AllZeroTable[1,])

# Minimum mismatches
THUL01Minnum_switches

THUL01Minnum_switchesTable = c("2018 THUL B","MM", "6", "699", "0 - 160", "0, 200 - 1000", "1, 10", "0", "1.00", "6", "6")

trackScores[nrow(trackScores) + 1,] = c(THUL01Minnum_switchesTable)

# Correct flower, minimum mismatches
THUL01CorrectflowerMinnum_switchesTable = c("2018 THUL B","CF MM", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA")
trackScores[nrow(trackScores) + 1,] = c(THUL01CorrectflowerMinnum_switchesTable)

# Add NARS-17 to table
# All zero
NARS17AllZero$Site = "2019 NARS D"
NARS17AllZero$Type = "PZ"

NARS17AllZeroTable = NARS17AllZero %>%
  select(Site,Type, num_unique_objects, num_detections, maxDisap, maxDist, runMean,num_switches, mota, mostly_tracked, num_tracks)

trackScores[nrow(trackScores) + 1,] = c(NARS17AllZeroTable[1,])

# Minimum mismatches
NARS17Minnum_switches

NARS17Minnum_switchesTable = c("2019 NARS D","MM", "13", "748", "0 - 160", "0, 300 - 1000", "1 - 160", "0", "1.00", "13", "7 - 13")

trackScores[nrow(trackScores) + 1,] = c(NARS17Minnum_switchesTable)

# Correct flower, minimum mismatches
NARS17CorrectflowerMinnum_switches = NARS17d1 %>% 
  filter(num_tracks == length(unique(nars17_gt$id_gt)))
NARS17CorrectflowerMinnum_switches = NARS17CorrectflowerMinnum_switches[NARS17CorrectflowerMinnum_switches$num_switches == min(NARS17CorrectflowerMinnum_switches$num_switches),]

NARS17CorrectflowerMinnum_switchesTable = c("2019 NARS D","CF MM", "13", "748", "0 - 60", "0, 300 - 1000", "1 - 160", "0", "1.00", "13", "13")

trackScores[nrow(trackScores) + 1,] = c(NARS17CorrectflowerMinnum_switchesTable)

# Add NARS-04 to table
# All zero
NARS04AllZero$Site = "2018 NARS A"
NARS04AllZero$Type = "PZ"

NARS04AllZeroTable = NARS04AllZero %>%
  select(Site,Type,num_unique_objects, num_detections, maxDisap, maxDist, runMean,num_switches, mota, mostly_tracked, num_tracks)

trackScores[nrow(trackScores) + 1,] = c(NARS04AllZeroTable[1,])

# Minimum mismatches
NARS04Minnum_switches$Site = "2018 NARS A"
NARS04Minnum_switches$Type = "MM"
NARS04Minnum_switchesTable = NARS04Minnum_switches %>%
  select(Site,Type,num_unique_objects, num_detections, maxDisap, maxDist, runMean,num_switches, mota, mostly_tracked, num_tracks)

trackScores[nrow(trackScores) + 1,] = c(NARS04Minnum_switchesTable[1,])

# Correct flower, minimum mismatches
NARS04CorrectflowerMinnum_switchesTable = c("2018 NARS A","CF MM", "32", "1724", "10", "500", "10", "19", "0.98898", "32", "32")
trackScores[nrow(trackScores) + 1,] = c(NARS04CorrectflowerMinnum_switchesTable)

# Add NYAA-04 to table
# All zero
NYAA04AllZero$Site = "2019 NYAA E"
NYAA04AllZero$Type = "PZ"

NYAA04AllZeroTable = NYAA04AllZero %>%
  select(Site,Type,num_unique_objects, num_detections, maxDisap, maxDist, runMean,num_switches, mota, mostly_tracked, num_tracks)

trackScores[nrow(trackScores) + 1,] = c(NYAA04AllZeroTable[1,])

# Minimum mismatches
NYAA04Minnum_switches

NYAA04Minnum_switchesTable = c("2019 NYAA E","MM", "21", "1762", "70 - 160", "600", "1", "17", "0.99035", "21", "21")

trackScores[nrow(trackScores) + 1,] = c(NYAA04Minnum_switchesTable)

# Correct flower, minimum mismatches
NYAA04CorrectflowerMinnum_switchesTable = c("2019 NYAA E","CF MM","NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA", "NA")

trackScores[nrow(trackScores) + 1,] = c(NYAA04CorrectflowerMinnum_switchesTable)

# Add NARS-13 to table
# All zero
NARS13AllZero$Site = "2019 NARS C"
NARS13AllZero$Type = "PZ"

NARS13AllZeroTable = NARS13AllZero %>%
  select(Site,Type,num_unique_objects, num_detections, maxDisap, maxDist, runMean,num_switches, mota, mostly_tracked, num_tracks)

trackScores[nrow(trackScores) + 1,] = c(NARS13AllZeroTable[1,])

# Minimum mismatches
NARS13Minnum_switches$Site = "2019 NARS C"
NARS13Minnum_switches$Type = "MM"
NARS13Minnum_switchesTable = NARS13Minnum_switches %>%
  select(Site,Type,num_unique_objects, num_detections, maxDisap, maxDist, runMean,num_switches, mota, mostly_tracked, num_tracks)

trackScores[nrow(trackScores) + 1,] = c(NARS13Minnum_switchesTable[1,])

# Correct flower, minimum mismatches
NARS13CorrectflowerMinnum_switchesTable = c("2019 NARS C","CF MM", "85", "4887", "10", "400", "10", "91", "0.98138", "85", "85")
trackScores[nrow(trackScores) + 1,] = c(NARS13CorrectflowerMinnum_switchesTable)

```

```{r}
write.table(trackScores, "U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/trackScores.csv", row.names = F, col.names = F, quote = F, sep=";")
```

```{r}

percDiff <- function(val1, val2) {
  pdiff = abs(val1-val2) / (abs((val1+val2)/2))

  return(pdiff*100)
}

trackScoresFig = trackScores %>% 
  select(Site, Tracks, Flowers, MOTA)

trackScoresFig$flowerError = percDiff(trackScores$Tracks, trackScores$Flowers)

```

# Camera codes:
# 2018 NARS-04: 2018 NARS A
# 2018 THUL-01: 2018 THUL B
# 2019 NARS-13: 2019 NARS C
# 2019 NARS-17: 2019 NARS D
# 2019 NYAA-04: 2019 NYAA E

####### /MMFix PARAMETER TEST ########