### 2023-01-20 src/R/tableGenerator.Rmd
Submitted revised edition of manuscript

---
title: "TableGenerator"
author: "Hjalte Mann"
date: '2022-05-02'
#output: pdf
pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(ggplot2)
library(dplyr)
library(stringr)
library(GGally)
library(plotly)
library(kableExtra)
library(gt)
install.packages("webshot2")
library(webshot2)

```

```{r echo = F}

trackScores = read.csv2("U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/trackScores.csv", header = F)

colnames(trackScores) = c('Site','Type', 'Flowers','Points', 'Max gap', 'Max distance', 'Running mean', 'Mismatches', 'MOTA', '>80% Accuracy','Tracks')

trackScores$MOTA = signif(as.numeric(trackScores$MOTA), digits = 3)

trackScores %>% 
  gt() %>%
  tab_spanner(
    label = "Tracking parameters",
    columns = c('Max gap', 'Max distance', 'Running mean'
    )
  )%>%
  tab_spanner(
    label = "Result",
    columns = c('Mismatches', 'MOTA', '>80% Accuracy', 'Tracks'
    )
  ) %>% 
   cols_align(
    align = "left"
  ) %>%
  tab_row_group(
    label = "2018 THUL B",
    rows = Site == "2018 THUL B"
  )  %>%
  tab_row_group(
    label = "2019 NARS D",
    rows = Site == "2019 NARS D"
  ) %>%
  tab_row_group(
    label = "2018 NARS A",
    rows = Site == "2018 NARS A"
  ) %>%
  tab_row_group(
    label = "2019 NYAA E",
    rows = Site == "2019 NYAA E"
  ) %>%
  tab_row_group(
    label = "2019 NARS C",
    rows = Site == "2019 NARS C"
  ) %>% tab_style(
    style = list(
      cell_fill(color = "#FFFFFF"),
      cell_text(color = "#FFFFFF")
      ),
    locations = cells_body(
      columns = Site
    )
  ) %>% 
  row_group_order(groups = c("2018 THUL B","2019 NARS D","2018 NARS A","2019 NYAA E", "2019 NARS C" )) %>%
  fmt_missing(
    columns = 1:11,
    missing_text = "-"
  ) %>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
      ),
    locations = cells_body(
      columns = Tracks,
      rows = c(1,2,4,6,9,11,12,15)
    )
  )%>% 
  text_transform(locations = cells_body(columns = Tracks), 
                 fn = function(x) {
                   x <- gsub("7 - 13", "7 - <span style=\"font-weight:bold;\">13</span>", x)
                 }) %>% 
  gtsave("tab_1_2.png", expand = 10)

```

```{r}

filterScores = read.csv2("U:/BITCue/Projekter/TrackingFlowers/testResults/MMFix/FilterScoresForTable.csv", header = T)

colnames(filterScores) = c('Image series','Flowers', 'Tracks','Mismatches', 'MOTA', 'Tracks ', 'Mismatches ', 'MOTA ')

filterScores$MOTA = signif(as.numeric(filterScores$MOTA), digits = 3)

filterScores %>% 
  gt() %>%
  tab_spanner(
    label = "Before filtering",
    columns = c('Tracks','Mismatches', 'MOTA'
    )
  )%>%
  tab_spanner(
    label = "After filtering",
    columns = c('Tracks ','Mismatches ', 'MOTA '
    )
  ) %>% 
   cols_align(
    align = "left"
  ) %>% 
  fmt_missing(
  columns = everything(),
  rows = everything(),
  missing_text = ""
) %>% tab_style(
    style = cell_borders(
      sides = c("top", "bottom"),
      color = "lightgrey",
      weight = px(2),
      style = "solid"
    ),
    locations = cells_body(
      columns = everything(),
      rows = 4
    )
  ) %>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
      ),
    locations = cells_body(
      columns = everything(),
      rows = 4
    )
  ) %>% 
  gtsave("tableFilterscores.png", expand = 10)
```

```{r}

seriesOverview = read.csv2("U:/BITCue/Projekter/TrackingFlowers/data/annotations/seriesOverview.csv", header = T)
colnames(seriesOverview) = c('Image series','Number of images', 'Number of flowers')
seriesOverview

seriesOverview  %>% 
  gt() %>%
  cols_align(
    align = "left") %>% 
    gtsave("U:/BITCue/Projekter/TrackingFlowers/src/R/seriesOverview.png", expand = 10)

```